---
title: "Using BaseR"
author: "Mike Keating"
format: pdf
editor: visual
---

## Task 1: Basic Vector Practice

### Question 1:

Create two vectors named pre and post. One vector corresponding to the pre measurements and one to the post measurements.. Create two vectors named pre and post. One vector corresponding to the pre measurements and one to the post measurements.

```{r}
pre <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

post <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

### Question 2:

Assign names to the vector elements using the paste() function. Note that names() can be overwritten by a character vector.

```{r}
names <- paste("Subject", 1:20, sep="_")
# Assign names
names(pre) <- names
names(post) <- names
```

### Question 3:

Calculate the change in blood pressure for each patient.

```{r}
# Change in blood pressure
diff_op <- pre - post
diff_op
```

### Question 4:

Calculate the average decrease in blood pressure across all patients.

```{r}
avg_change_all <- mean(diff_op)

```

### Question 5

Determine which patients experienced a decrease in blood pressure after treatment (a positive change). Use the which() function to just return the indices (and names) associated with this type of change.

```{r}
index_patients_decreased_bp <- which(diff_op > 0)



```

### Question 6

Subset the vector of differences to only return those that have a positive change

```{r}
patients_decreased_bp <- diff_op[index_patients_decreased_bp]
```

### Question 7

Calculate the average decrease in blood pressure for those where the blood pressure decreased (positive change).

```{r}
mean_patients_decreased_bp <- mean(patients_decreased_bp)
paste("The average decrease in blood pressure for those where the blood pressure decreased was: ", mean_patients_decreased_bp)
```

## Task 2: Basic Data Frame Practice

```{r}
df <- data.frame("Paiteint" = names, "pre" = pre, "post"=post, "diff_op"=diff_op)
```
